The Economics of Shotgun Marriage

### 2020-10-02 slides-may-20/slides.tex
Current stage

% REMOVE SLIDE COUNT

@@

\begin{frame}
\frametitle{``America's strongest anti-poverty weapon''?}
\begin{itemize}
\item Married two-parent families \ : \ desirable outcomes for parents and children%$\genfrac{}{}{0pt}{1}{\Leftarrow}{\Rightarrow}$
\item Pro-marriage policies: child support, tax brackets, divorce regulations, etc
\item Can economic policies create successful marriages? Love vs money.
\end{itemize}
\end{frame}

\begin{frame}

\frametitle{Kids-First Marriages are Less Stable}
\begin{enumerate}
\item{\textbf{Shotgun marriage}} --- couple marrying when woman gets pregnant
\item{\textbf{Kids-first marriage}} --- child \textit{before or in the year} of marriage
\end{enumerate}
\begin{itemize}
\item Around 1/4 marriages in the US today are kids-first
\item They have much higher divorce rates, especially among college graduates
\item Marginal in economic sense: could be sensitive to policies 
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Roadmap}
\begin{enumerate}
\item Document the shotgun marriage and its outcomes empirically
\item Use these empirical relations to estimate a model capturing:
\begin{itemize}
\item dynamic bargaining and limited commitment in marriage
\item lifecycle, income and savings
\item fertility and child expenditures
\end{itemize}
\item Look at what drives decisions to marry and the differences in divorce
\begin{itemize}
\item what explains the education gradient?
\item can policies fix the difference or reduce the share?
\item what is the value of marriage in general?
\end{itemize}
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Summary of Results}
\begin{itemize}
\item 23\% of women had the first child before their first marriage (kids first)
\item They are 1.8 times more likely to be divorced than ``marriage first''
%\item College grads: 2.8 times more likely to be divorced
%\begin{itemize}
%\item though smaller kids first share (just 10\%)
%\end{itemize}
\item My structural model corrects selection and data imperfections
%\item Model results:
%\begin{itemize}
%\item Social stigma + threat of raising the child alone explains most of the difference
%\item Child support $\Rightarrow$ more single mothers, little effect on marriage decisions
%\item Being able to control fertility helps everyone
%\end{itemize}
\item Model answers: Love $>$ Money. In particular:
\begin{itemize}
\item Threat of raising the child alone explains most of the difference
\item Regulating marriages by policies is both hard and inefficient
\item Easier control over fertility fixes shotgun marriages and has bigger return
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Literature}
\begin{itemize}
\item Policy literature:
\begin{itemize}
\item Empirical: Alessina, Guiliano, 2005;  Tannenbaum, 2020; Rossin-Slater, 2017
\item Structural: Forester, 2020; Brown, Flinn, 2011
\item Focus on particular policies, ignore interactions, limited quantitative power
\end{itemize}
\item Lifecycle literature:
\begin{itemize}
\item Dynamic limited commitment for marriage and divorce: Mazzocco, 2007; Voena, 2015; Low et al., 2018; Shephard, 2019; Blasutto, 2020
\item Fertility during lifecycle: Sommer, 2015, Ejrnæs, Jørgensen, 2018
\item Link between divorce and fertility is missing
\end{itemize}
\item Empirics on shotgun marriages (demography and sociology):
\begin{itemize}
\item Gibson-Davis et al (2016), Guzzo (2009),  Nuevo-Chiquero (2014)
\item Mostly descriptive
\end{itemize}
\end{itemize}
%
%\begin{itemize}
%\item Empirical literature documents the scope of shotgun marriage around the world:
%\begin{itemize}
%\item Gibson-Davis et al (2016), Guzzo (2009),  Nuevo-Chiquero (2014)
%\item Mostly descriptive
%\end{itemize}
%\item Dynamic limited commitment models of divorce and marriage:
%\begin{itemize}
%\item Marcet, Marimon (2019), Mazzocco (2007) --- setting up the framework
%\item Voena (2015), Shephard (2019); Lafortune, Low  (2019) --- applied questions
%\item Mostly ignore or simplify interaction with fertility
%\end{itemize}
%\item Unplanned and teenage pregnancy literature:
%\begin{itemize}
%\item Ejrnæs, Jørgensen (2018) --- lifecycle model
%\item Card, Wise (1978) to Ashcraft et al (2013) --- lots of empirical papers
%\item Treat household structure as fixed or exogenous
%\end{itemize}
%\end{itemize}
\end{frame}

\section{Data}
\begin{frame}
\frametitle{Sources: ACS 2009--2017 and others}
\begin{itemize}
\item Main exercises and model estimation are done on American Community Survey
%\item Huge sample size and enough information to show many things 
\item Individual marriage and fertility history are observed partially
\begin{itemize}
\item Only the most recent marriage date --- restrict to ``married once'' when matters
\item Pick age $\in [21,40]$ to have good information on fertility
\end{itemize}
\item Smaller datasets with better histories for robustness and extra evidence:
\begin{itemize}
\item Survey of Income and Program Participation (SIPP)
\item More: NSFG and NSFH
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[label=dt_graph]
\frametitle{Kids First Women Are Divorced More Often: Just Look}
\begin{center}
\includegraphics[scale=0.75]{div_5y_by_dt.pdf}
\end{center}
\vspace{-1cm}
\hyperlink{dt_graph_educ}{\beamerbutton{by education}} 
\end{frame}

\begin{frame}
\frametitle{Kids First Women Are Divorced More Often: Quantification}
\framesubtitle{ACS 2009--2017, Women 21--40, married once}
\begin{center}
\begin{tabular}{l r r r r }
\hline
& \multicolumn{2}{c}{share of divorced if ... }&  \\
&  marriage-first & kids-first & (share of kids-first) &  \\\hline
\multicolumn{5}{l}{\textbf{All sample}} \\\hline
\textit{Cross-sectional share of divorced} &  9.9 & 18.1 & (26.1) \\
\textit{Divorced 5 years after marriage} &  5.1 & 14.3  & (24.6) \\
\textit{Divorced 10 years after marriage} & 10.4 & 22.8 & (24.4) \\\hline
\multicolumn{5}{l}{\textbf{Cross-sectional share of divorced for subsamples}} \\\hline
\textit{High school only} &  12.8 & 17.3 & (37.0) \\
\textit{Some college} & 14.3 & 21.0 & (31.3) \\
\textit{College or more} &   5.3 & 14.8 & (11.9) \\\hline
\textit{First birth before 25} & 15.5 & 19.9 & (41.8) \\
\textit{First birth at 25 or later} &  6.2 & 10.9 & (10.7)  \\\hline
\end{tabular}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Kids First Women Are Divorced More Often: Controls}
\begin{center}
%\vspace{-0.5cm}
\begin{tabular}{l r r r }
\multicolumn{4}{c}{\textit{Regression equation:} $\text{Divorced}_i = \Delta \cdot \text{kids-first}_i + \text{Controls}_i + \varepsilon_i$} \\\hline
\hline
& \multicolumn{2}{c}{Estimates}  \\
& Difference ($\Delta$) & (standard error)  \\\hline
\multicolumn{4}{l}{\textbf{All sample, regression}} \\\hline
\textit{No controls (raw difference) }& 8.2 & (0.1)  \\
\textit{Demographic controls }& 5.7 & (0.1)  \\
\textit{Duration controls} &  5.8 & (0.2)  \\
\textit{Demographic + duration} &  4.8 & (0.2)  \\
\textit{Demographic + duration + income} & 4.1 & (0.3)  \\\hline
\multicolumn{4}{l}{\textbf{All sample, propensity score matching}} \\\hline
\textit{Demographic controls} & 5.0 & (0.3)  \\
\textit{Demographic controls + income} & 4.9 & (0.2)  \\\hline
\multicolumn{4}{p{0.825\linewidth}}{\footnotesize \textit{Controls:} demographics: FE for age $\times$ education, race, state, survey year. Duration: FE for age $\times$ age of the first marriage and age $\times$ age of the first birth.  Income: 3rd degree polynomial of log gross income (if observed).}\\\hline\hline
\end{tabular}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Kids First Women Are Divorced More Often: Extra Evidence from SIPP}
\begin{center}
\begin{tabular}{l r r }
\hline
\multicolumn{3}{l}{\textbf{Including remarried people...}} \\\hline
& \multicolumn{2}{c}{share of divorced if ... }  \\
&  marriage-first & kids-first   \\\hline
\textit{Ever divorced} & 34.4 & 48.0  \\
\textit{Divorced if married once} &  12.4 & 19.3 \\
\textit{Remarried} &  22.0 & 28.7 \\\hline
\hline
\multicolumn{3}{l}{\textbf{Separating step and own children (2+ kids only)}} \\\hline
& share in sample & share of divorced   \\\hline
{Marriage-first} & 72.1 & 10.5 \\
{Kids-first-own} &  16.4 & 14.6 \\
{Kids-first-other} &  11.4 & 27.2 \\\hline
\multicolumn{3}{p{0.7\linewidth}}{\footnotesize \textit{Notes.} SIPP data, the cross-section of Wave 1, 2014, women 21+. Numbers are percents.}\\\hline\hline
\end{tabular}
\end{center}
\end{frame}

%\begin{frame}
%\frametitle{Kids First Women Are Divorced More Often: Robustness \& Extra}
%\begin{itemize}
%\item SIPP exercise: smaller sample but full marital history and fertility history
%\begin{itemize}
%\item Similar ratios with larger magnitudes when include remarried people
%\end{itemize}
%\item Duration model: NSFH, historic data from 80s
%\begin{itemize}
%\item Hazard ratio of 1.45, college: 2.23
%\end{itemize}
%\item More details: NSFG, 2015--2017 data
%\begin{itemize}
%\item KF: 30\% true shotgun, 50\% marry after birth, 20\% marry new partner
%\item True shotgun have the highest share of divorced
%\end{itemize}
%\end{itemize}
%\end{frame}
%
%
%\begin{frame}
%\frametitle{Need For The Model}
%\begin{enumerate}
%\item Observables provide limited information. Missing things: relationship quality, potential partners, future divorces
%\item Data are not perfect. Limited histories, unobservable income if out of labor force, step or own children
%\end{enumerate}
%\end{frame}

@@

\begin{frame}
\frametitle{Modeling Shotgun Marriage}
\begin{center}
\vspace{-0.6cm}
\begin{tikzpicture}[every text node part/.style={align=center}, scale=0.75]
   % Place nodes
   \node [block] (1) {\large \textbf{People}  \\[-0.5ex]  \large  \textbf{Randomly Meet}};
   \node [block, below left = 1.5 cm and 0.0 cm of 1] (2) {\textit{Nash} \\[-0.5ex] \textit{Bargaining}};   
   \node [block, below right = 1.5 cm and 0.0 cm of 1] (3) {\textit{Nash} \\[-0.5ex] \textit{Bargaining}};

   \node [block, below = 1.5 cm of 2] (5) {\textbf{Kids First} \\[-0.5ex] \textbf{Couple} ($ck$)};
   \node [block, below = 0.8 cm of 3] (7) {No Kids \\[-0.5ex] Couple ($c$)};
   \node [block, below = 0.8 cm of 7] (8) {\textbf{Mariage First} \\[-0.5ex] \textbf{Couple} ($ck$)};
   \draw[->] (1) to [bend left = 20] node [left] {\scriptsize Unplanned \\[-0.75ex] \scriptsize pregnancy} (2);
   \draw[->] (1) to [bend right = 20] node [right] {\scriptsize No \\[-0.75ex] \scriptsize pregnancy} (3);
   \draw[->] (2) to node [left] {\scriptsize Agree} node [right] {\scriptsize \textbf{Shotgun} \\[-0.75ex]  \scriptsize \textbf{marriage}} (5);
   \draw[->] (3) to node [right] {\scriptsize Agree} node [left]  {\scriptsize \textbf{Regular} \\[-0.75ex]  \scriptsize \textbf{marriage}} (7);
   \draw[->] (7) to node [right] {\scriptsize Kids by choice} (8);   
    \node [block, above left   = 0.8 cm and -0.75 cm of 2] (4) {\,Single male \\[-0.5ex] ($m,s$)\,\,};
    \node [block, above left   = 1.8 cm and -0.75 cm of 2] (4-sm) {\,Single mom \\[-0.5ex]  ($f,sk$)};
    %\node [block, left = 0.2 cm of 4-sm] (4-nsm) {\,Single female \\[-0.5ex]  ($f,sk$)};
	\node [block, above right   = 0.8 cm and -0.75 cm of 3] (6) {\,\,Single male\,\,\\[-0.5ex] ($m,s$)};
    \node [block, above right   = 1.8 cm and -0.75 cm of 3] (6-f) {Single female\\[-0.5ex] ($f,s$)\,};   
    \draw[->] (2.180) to [bend left = 30] node [left] {\scriptsize Disagree} (4.270);
    \draw[->,thick,dotted] (4-sm.20) to [bend left = 10] node [above right=-0.2cm and 0.7cm] {\scriptsize Costly abortion \\[-0.75ex] \scriptsize (if possible)} (6-f.50);
    \draw[->] (3.0) to [bend right = 30] node [right] {\scriptsize Disagree} (6.270);
     \draw[->] (6.180) to [bend left = 30] node [above] {\scriptsize Back}  (1.-12);
   \draw[->] (6-f.180) to [bend right = 30] node [above] {\scriptsize Back} (1.12);
   \draw[->,thick,dashed] (4-sm.90) to [bend left = 30] node [above] {\scriptsize Worse marriage prospectives}  (1.168);
   \draw[->] (4.0) to [bend right = 30] node [above] {\scriptsize Back} (1.192);

@@

\begin{frame}
\frametitle{Value Function: Female, Single}
\framesubtitle{$p^m_t$ --- probability of meeting, $p^p_t$ --- probability of unplanned pregnancy, $m_t$ --- agree to marry}
\begin{align*}
V^{f,s}_t(a,w) = \max\limits_{c,a'} \Bigg\{ \frac{c^{1-\sigma}}{1-\sigma} + \beta \cdot \E_t \Big\{  & 
%& \\ \text{ (did not meet anyone): \ } & 
(1-p^{\text{m}}_t)\cdot V_{t+1}^{f,s}(a',w') \\
\text{ \footnotesize (met and not pregnant): \ }& +  p^{\text{m}}_t\cdot (1- p^{p}_t) \cdot \left[ m^{np}_t \cdot V^{f,c}_{t+1}  + (1-m^{np}_t)\cdot  V^{f,s}_{t+1}\right] \\
\text{ \footnotesize (met and pregnant): \ } & + p^{\text{m}}_t \cdot p^{\text{p}}_t\cdot   \left[ m^p_t \cdot V^{f,ck}_{t+1}  + (1-m^p_t)\cdot  V^{f,\text{preg}}_{t+1}\right] \Big\} \Bigg\}, \\
\text{ \footnotesize (uncertain abortion access): } &  V^{f,\text{preg}}_{t+1} = p^{\text{access}} \cdot \max\{V^{f,sk}_{t+1},V^{f,s}_{t+1} - \phi_a\} \\
   & \hspace{1cm} + (1-p^{\text{access}})\cdot V^{f,sk}_{t+1},\\
\text{ \footnotesize (budget constraint): \ } & c + a' = R\cdot a  + w , \\
\text{ \footnotesize (wage evolution): \ } & \log w = z + \text{Trend}_t, \ \ z' = z + \varepsilon^{f}.
\end{align*}
\end{frame}

\begin{frame}
\frametitle{Value Function: Couple, No Children}
\framesubtitle{$k_t$ --- give a birth, $d_t$ --- get a divorce, $\tilde{V}$ --- values at divorce}
\begin{align*}
V^{c}_t(a,w_f,w_m,\psi,\theta) = \max\limits_{c_f,c_m,a',k} \Bigg\{ & \theta^f \cdot \frac{c_f^{1-\sigma}}{{1-\sigma}}  + \theta^m \cdot \frac{c_m^{1-\sigma}}{1-\sigma}  + \psi + \beta \cdot \E_t \Big\{   \\
\text{ \footnotesize (stay, no birth): \ } & + (1-d_t) \cdot (1-k_t) \cdot  \left[ \theta^f \cdot  V^{f,c}_{t+1} + \theta^m \cdot V^{m,c}_{t+1}  \right] \\
\text{ \footnotesize (stay, give a birth): \ } &  + (1-d_t) \cdot k_t \cdot  \left[ \theta^f \cdot  V^{f,ck}_{t+1} + \theta^m \cdot V^{m,ck}_{t+1}  \right]  \\
\text{ \footnotesize (divorce): \ } & + d_t  \left[ \theta^f \cdot  \tilde{V}^{f,s}_{t+1} + \theta^m \cdot \tilde{V}^{m,s}_{t+1}  \right]  \Big\} \Bigg\}\\
\text{ \footnotesize (budget constraint, rts): \ } &{\normalsize \left[c_f^{1+\rho} + c_m^{1+\rho}\right]^{\frac1{1+\rho}} + a' = R\cdot a + w_f + w_m}\\
\text{ \footnotesize (match quality evolution): \ } &{\normalsize \psi' = \psi + \varepsilon^{\psi}}\\
\text{ \footnotesize \textbf{(participation constraints)}: \ } &{\normalsize V^{f,c\bullet}_{t+1} (\cdots,\theta') \geq \tilde{V}^{f,s}_{t+1}, \ \ V^{m,c\bullet}_{t+1} (\cdots,\theta') \geq \tilde{V}^{m,s}_{t+1}.}\\
\end{align*}
\end{frame}

\begin{frame}
\frametitle{Value Function: Couple and Child}
\framesubtitle{$d_t$ --- get a divorce, $\tilde{V}$ --- values at divorce}
\begin{align*}
V^{ck}_t(a,w_f,w_m,\psi,\theta) = \max\limits_{c_f,c_m,x,l_f,a'} \Bigg\{ & \theta^f \cdot \frac{c_f^{1-\sigma}}{{1-\sigma}}  + \theta^m \cdot \frac{c_m^{1-\sigma}}{1-\sigma}  + \psi + \\ \text{ \footnotesize (child is a public good): \ }  & \phi + \alpha \cdot \frac{Q^{1-\sigma}}{1-\sigma} +  \beta \cdot \E_t \Big\{   \\
\text{ \footnotesize (stay): \ } & + (1-d_t) \cdot  \left[ \theta^f \cdot  V^{f,ck}_{t+1} + \theta^m \cdot V^{m,ck}_{t+1}  \right] \\
\text{ \footnotesize (divorce): \ } & + d_t \cdot \left[ \theta^f \cdot  \tilde V^{f,sk}_{t+1} + \theta^m \cdot \tilde V^{m,s}_{t+1}  \right]  \Big\} \Bigg\}\\
\text{ \footnotesize (budget constraint): \ } &{\normalsize \left[c_f^{1+\rho} + c_m^{1+\rho}\right]^{\frac1{1+\rho}} + x + a' = R\cdot a + w_f\cdot l_f + w_m}\\
\text{ \footnotesize (child consumption flow): \ } &{\normalsize Q = \left[x^\lambda + \kappa \cdot (1-l_f)^\lambda\right]^{\frac1\lambda}, \ \ \ Q\geq \bar{Q}} \text{ \footnotesize  (subsistence constraint) } \\
\text{ \footnotesize (female skills $\downarrow$ if not working): \ } &{\normalsize \log w_f = z_f + \text{Trend}_{f,t}, \ \ z'_f = z_f + \varepsilon^f - \delta(l^f)}\\
\end{align*}
\end{frame}

>>>>

@@

\begin{frame}
\frametitle{Two-Parent Family Is A Choice}

\textbf{Shotgun marriage} --- marriage that happens soon after a woman gets pregnant.
% many, much less stable, worse among the educated. pt is economics how they get formed, how do the
\begin{itemize}
\item Around 1/4 couples in the US have kids \textbf{before} or \textbf{at the year} of their marriage
\item These marriages are significantly less stable, especially among \textbf{educated people} % marginal marriage
\item Both social norms and economics are expected to drive this
%\item Especially less stable among \textbf{more educated and adult}%(college: 6\% single mothers, 4\% in shotgun marriages)
\item Overall, they contribute to creation of single mothers and blended families
% college grads: 4% in KF marriages and 4% are single mothers
% high school grads: 11% in KF marriages and 20% are single mothers
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Roadmap}
\begin{enumerate}
\item Document the shotgun marriage and its outcomes empirically
\item Use these empirical relations to estimate a model capturing:
\begin{itemize}
\item limited commitment in marriage
\item lifecycle, income and savings
\item fertility and child expenditures
\end{itemize}
\item Look at what drives the difference in marriage performance
\begin{itemize}
\item what explains the education gradient?
\item should policies bother to fix the difference?
\item what is the value of marriage in general?
\end{itemize}
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Summary of Results}
\begin{itemize}
\item 23\% of women had the first child before their first marriage (kids first)
\item They are 1.8 times more likely to be divorced than ``marriage first''
%\item College grads: 2.8 times more likely to be divorced
%\begin{itemize}
%\item though smaller kids first share (just 10\%)
%\end{itemize}
\item My structural model corrects selection and data imperfections
\item Model answers:
\begin{itemize}
%\item Causal content: kids first people divorce X times more
\item Social stigma explains 40\%--50\% of the divorce differences for college people
\item Expensive childcare explains the rest
\item Stigma matters less for high school graduates
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Literature}
\begin{itemize}
\item Empirical literature documents the scope of shotgun marriage around the world:
\begin{itemize}
\item Gibson-Davis et al (2016), Guzzo (2009),  Nuevo-Chiquero (2014)
\item Mostly descriptive
\end{itemize}
\item Dynamic limited commitment models of divorce and marriage:
\begin{itemize}
\item Marcet, Marimon (2019), Mazzocco (2007) --- setting up the framework
\item Voena (2015), Shephard (2019); Lafortune, Low  (2019) --- applied questions
\item Mostly ignore or simplify interaction with fertility
\end{itemize}
\item Unplanned and teenage pregnancy literature:
\begin{itemize}
\item Ejrnæs, Jørgensen (2018) --- lifecycle model
\item Card, Wise (1978) to Ashcraft et al (2013) --- lots of empirical papers
\item Treat household structure as fixed or exogenous
\end{itemize}
\end{itemize}
\end{frame}

\section{Data}
\begin{frame}
\frametitle{Sources: ACS 2009--2015 and others}
\begin{itemize}
\item Main exercises and model estimation are done on American Community Survey
\item Huge sample size and enough information to show many things 
\item Individual marriage and fertility history are observed partially
\begin{itemize}
\item Only the most recent marriage date --- restrict to ``married once'' when matters
\item Pick age $\in [21,40]$ to have good information on fertility
\end{itemize}
\item Smaller datasets with better histories for robustness and extra evidence:
\begin{itemize}
\item SIPP, NSFG, NSFH --- all have similar patterns
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[label=dt_graph]
\frametitle{Kids First Women Are Divorced More Often: Key Graph}
\begin{center}
\includegraphics[scale=0.75]{div_5y_by_dt.pdf}
\end{center}
\vspace{-1cm}
\hyperlink{dt_graph_educ}{\beamerbutton{by education}} 
\end{frame}

\begin{frame}
\frametitle{Kids First Women Are Divorced More Often: Quantification}
\begin{center}
\begin{tabular}{l c c c c c}\hline
& &  \multicolumn{2}{c}{\small \% divorced if ...} \\
& \% KF & MF & KF & \% diff &  Ratio \\\hline
\multicolumn{6}{c}{\textit{Cross-Section: All Females 21--40 married once}}\\
All females & 23.2 &  10.2 & 18.1 & 7.9 & 1.77  \\
College & 10.0 & 5.4 & 14.8 & 9.4 & 2.76 \\
High school & 34.3 & 13.9 & 17.3 & 3.3 & 1.24 \\\hline
\multicolumn{6}{c}{\textit{Controlling strategies}}\\\hline
%\multicolumn{2}{l}{College 5 y after marriage} & 5.1 & 14.3 & 9.2 & 2.79 \\
\multicolumn{4}{l}{Regression with controls}   & 7.0 & 1.69 \\
\multicolumn{4}{l}{Regression with controls + income}   & 6.1 & 1.60 \\\hline
\multicolumn{6}{p{0.7\linewidth}}{\footnotesize Regression: $\text{Divorced}_i = \Delta \cdot \text{KF}_i + X_i\gamma + \varepsilon_i$ (if $\text{KF}_i = 1$ or $\text{MF}_i = 1$)}\\
\multicolumn{6}{p{0.7\linewidth}}{\footnotesize Controls are two-way FE for interactions of age with education, time since marriage, time since birth, and one-way FE for geography, race, survey year. 
Income is 3rd degree polynomial of log per-hour reported income for working $\geq 10$ hours per week and earning $\geq \$3000$ per year.
}\\
\multicolumn{6}{p{0.7\linewidth}}{\footnotesize Data: American Community Survey 2009--2016.}\\\hline
\end{tabular}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Kids First Women Are Divorced More Often: Robustness \& Extra}
\begin{itemize}
\item SIPP exercise: smaller sample but full marital history and fertility history
\begin{itemize}
\item Similar ratios with larger magnitudes when include remarried people
\end{itemize}
\item Duration model: NSFH, historic data from 80s
\begin{itemize}
\item Hazard ratio of 1.45, college: 2.23
\end{itemize}
\item More details: NSFG, 2015--2017 data
\begin{itemize}
\item KF: 30\% true shotgun, 50\% marry after birth, 20\% marry new partner
\item True shotgun have the highest share of divorced
\end{itemize}
\end{itemize}
\end{frame}
%
%
%\begin{frame}
%\frametitle{Need For The Model}
%\begin{enumerate}
%\item Observables provide limited information. Missing things: relationship quality, potential partners, future divorces
%\item Data are not perfect. Limited histories, unobservable income if out of labor force, step or own children
%\end{enumerate}
%\end{frame}

@@

\begin{frame}
\frametitle{Modeling Shotgun Marriage}
\begin{center}
\vspace{-0.2cm}
\begin{tikzpicture}[every text node part/.style={align=center}, scale=0.75]
   % Place nodes
   \node [block] (1) {\large \textbf{People}  \\[-0.5ex]  \large  \textbf{Randomly Meet}};
   \node [block, below left = 1.5 cm and 0.0 cm of 1] (2) {\textit{Nash} \\[-0.5ex] \textit{Bargaining}};   
   \node [block, below right = 1.5 cm and 0.0 cm of 1] (3) {\textit{Nash} \\[-0.5ex] \textit{Bargaining}};

   \node [block, below = 1.5 cm of 2] (5) {\textbf{Kids First} \\[-0.5ex] \textbf{Couple} ($ck$)};
   \node [block, below = 0.8 cm of 3] (7) {No Kids \\[-0.5ex] Couple ($c$)};
   \node [block, below = 0.8 cm of 7] (8) {\textbf{Mariage First} \\[-0.5ex] \textbf{Couple} ($ck$)};
   \draw[->] (1) to [bend left = 20] node [left] {\scriptsize Unplanned \\[-0.75ex] \scriptsize pregnancy} (2);
   \draw[->] (1) to [bend right = 20] node [right] {\scriptsize No \\[-0.75ex] \scriptsize pregnancy} (3);
   \draw[->] (2) to node [left] {\scriptsize Agree} node [right] {\scriptsize \textbf{Shotgun} \\[-0.75ex]  \scriptsize \textbf{marriage}} (5);
   \draw[->] (3) to node [right] {\scriptsize Agree} node [left]  {\scriptsize \textbf{Regular} \\[-0.75ex]  \scriptsize \textbf{marriage}} (7);
   \draw[->] (7) to node [right] {\scriptsize Kids by choice} (8);   
    \node [block, above left   = 0.8 cm and -0.75 cm of 2] (4) {\,Single male \\[-0.5ex] ($m,s$)\,\,};
    \node [block, above left   = 1.8 cm and -0.75 cm of 2] (4-sm) {\,Single mom \\[-0.5ex]  ($f,sk$)};
	\node [block, above right   = 0.8 cm and -0.75 cm of 3] (6) {\,\,Single male\,\,\\[-0.5ex] ($m,s$)};
    \node [block, above right   = 1.8 cm and -0.75 cm of 3] (6-f) {Single female\\[-0.5ex] ($f,s$)\,};   
    \draw[->] (2.180) to [bend left = 30] node [left] {\scriptsize Disagree} (4.270);
    \draw[->] (3.0) to [bend right = 30] node [right] {\scriptsize Disagree} (6.270);
     \draw[->] (6.180) to [bend left = 30] node [above] {\scriptsize Back}  (1.-12);
   \draw[->] (6-f.180) to [bend right = 30] node [above] {\scriptsize Back} (1.12);
   \draw[->,thick,dashed] (4-sm.90) to [bend left = 30] node [above] {\scriptsize Worse marriage prospectives}  (1.168);
   \draw[->] (4.0) to [bend right = 30] node [above] {\scriptsize Back} (1.192);

@@

\begin{frame}
\frametitle{Value Function: Female, Single}
\framesubtitle{$p^m_t$ --- probability of meeting, $p^p_t$ --- probability of unplanned pregnancy, $m_t$ --- agree to marry}
\begin{align*}
V^{f,s}_t(a,w) = \max\limits_{c,a'} \Bigg\{ \frac{c^{1-\sigma}}{1-\sigma} + \beta \cdot \E_t \Big\{  & 
%& \\ \text{ (did not meet anyone): \ } & 
(1-p^{\text{m}}_t)\cdot V_{t+1}^{f,s}(a',w') \\
\text{ \footnotesize (met and not pregnant): \ }& +  p^{\text{m}}_t\cdot (1- p^{p}_t) \cdot \left[ m^{np}_t \cdot V^{f,c}_{t+1}  + (1-m^{np}_t)\cdot  V^{f,s}_{t+1}\right] \\
\text{ \footnotesize (met and pregnant): \ } & + p^{\text{m}}_t \cdot p^{\text{p}}_t\cdot   \left[ m^p_t \cdot V^{f,ck}_{t+1}  + (1-m^p_t)\cdot  V^{f,sk}_{t+1}\right] \Big\} \Bigg\} \\
\text{ \footnotesize (budget constraint): \ } & c + a' = R\cdot a  + w , \\
\text{ \footnotesize (wage evolution): \ } & \log w = z + \text{Trend}_t, \ \ z' = z + \varepsilon^{f}.
\end{align*}
\end{frame}

\begin{frame}
\frametitle{Value Function: Couple, No Children}
\framesubtitle{$k_t$ --- give a birth, $d_t$ --- get a divorce, $\tilde{V}$ --- values at divorce}
\begin{align*}
V^{c}_t(a,w_f,w_m,\psi,\theta) = \max\limits_{c_f,c_m,a',k} \Bigg\{ & \theta^f \cdot \frac{c_f^{1-\sigma}}{{1-\sigma}}  + \theta^m \cdot \frac{c_m^{1-\sigma}}{1-\sigma}  + \psi + \beta \cdot \E_t \Big\{   \\
\text{ \footnotesize (stay, no birth): \ } & + (1-d_t) \cdot (1-k_t) \cdot  \left[ \theta^f \cdot  V^{f,c}_{t+1} + \theta^m \cdot V^{m,c}_{t+1}  \right] \\
\text{ \footnotesize (stay, give a birth): \ } &  + (1-d_t) \cdot k_t \cdot  \left[ \theta^f \cdot  V^{f,ck}_{t+1} + \theta^m \cdot V^{m,ck}_{t+1}  \right]  \\
\text{ \footnotesize (divorce): \ } & + d_t  \left[ \theta^f \cdot  \tilde{V}^{f,s}_{t+1} + \theta^m \cdot \tilde{V}^{m,s}_{t+1}  \right]  \Big\} \Bigg\}\\
\text{ \footnotesize (budget constraint): \ } &{\normalsize \left[c_f^{1+\rho} + c_m^{1+\rho}\right]^{\frac1{1+\rho}} + a' = R\cdot a + w_f + w_m}\\
\text{ \footnotesize (match quality evolution): \ } &{\normalsize \psi' = \psi + \varepsilon^{\psi}}\\
\text{ \footnotesize \textbf{(participation constraints)}: \ } &{\normalsize V^{f,c\bullet}_{t+1} (\cdots,\theta') \geq \tilde{V}^{f,s}_{t+1}, \ \ V^{m,c\bullet}_{t+1} (\cdots,\theta') \geq \tilde{V}^{m,s}_{t+1}.}\\
\end{align*}
\end{frame}

\begin{frame}
\frametitle{Value Function: Couple and Child}
\framesubtitle{$d_t$ --- get a divorce, $\tilde{V}$ --- values at divorce}
\begin{align*}
V^{ck}_t(a,w_f,w_m,\psi,\theta) = \max\limits_{c_f,c_m,x,l_f,a'} \Bigg\{ & \theta^f \cdot \frac{c_f^{1-\sigma}}{{1-\sigma}}  + \theta^m \cdot \frac{c_m^{1-\sigma}}{1-\sigma}  + \psi + \\ \text{ \footnotesize (child is a public good): \ }  & \phi + \alpha \cdot \frac{Q^{1-\sigma}}{1-\sigma} +  \beta \cdot \E_t \Big\{   \\
\text{ \footnotesize (stay): \ } & + (1-d_t) \cdot  \left[ \theta^f \cdot  V^{f,ck}_{t+1} + \theta^m \cdot V^{m,ck}_{t+1}  \right] \\
\text{ \footnotesize (divorce): \ } & + d_t  \left[ \theta^f \cdot  \tilde V^{f,sk}_{t+1} + \theta^m \cdot \tilde V^{m,s}_{t+1}  \right]  \Big\} \Bigg\}\\
\text{ \footnotesize (budget constraint): \ } &{\normalsize \left[c_f^{1+\rho} + c_m^{1+\rho}\right]^{\frac1{1+\rho}} + x + a' = R\cdot a + w_f\cdot l_f + w_m}\\
\text{ \footnotesize (child quality flow): \ } &{\normalsize Q = \left[x^\lambda + \kappa \cdot (1-l_f)^\lambda\right]^{\frac1\lambda}, \ \ \ Q\geq \bar{Q}} \text{ \footnotesize  (subsistence constraint) } \\
\text{ \footnotesize (female skills depreciate): \ } &{\normalsize \log w_f = z_f + \text{Trend}_{f,t}, \ \ z'_f = z_f + \varepsilon^f - \delta(l^f)}\\
\end{align*}
\end{frame}

\begin{frame}
\frametitle{Renegotiation and Divorce}
\framesubtitle{Are generated by the participation constraints}
\begin{itemize}
\item Bargaining weights $\theta$ adjust to satisfy both participation constraints.
\item If no adjustment is possible, the couple divorces
\item Assets are split evenly upon divorce
\item If couple with a child divorces wife becomes a single mother
\item Child support: adjustment in permanent income component $z$
\end{itemize}
\end{frame}
